Pretty darn good control: when are approximate solutions better than approximate models

### 2023-06-16 manuscript/manuscript.Rmd
mini
improved the methods section a bit

\textbf{The Environment.}
As seen in Fig. **TBD CONCEPTUAL**, we consider four different models of the ecosystem with which the fishery interacts, each of this is called an *environment*, alluding to the environment in an RL setting.
To recap, these four environments are a single-species growth model \eqref{eq:may77} for a harvested species; a three-species model \eqref{eq:3d model stoch} with a single harvested species; the same three-species model but with *two* harvested species; and, finally, a three-species model with a time-varying parameter and two harvested species.

The policies explored were functions from states $N_t$ to a single number, $M_X(N_t) = \pi_\theta(N_t)$, in the single fishery case.
In the two-fishery case, policies lead to two-dimensional actions: $(M_X(N_t), M_Y(N_t)) = \pi_\theta(N_t)$.
Each episode in the training and in the evaluation stages was at most 200 time steps long.
Episodes were ended early in case of a ‘‘near-extinction event,’’ a moment in which any of the populations dipped below a threshold value, i.e.,
\begin{align}
  \label{eq:thresh}
  X_t \leq X_{\text{thresh.}},\quad
  Y_t \leq Y_{\text{thresh.}},\quad
  \text{or,}\quad
  Z \leq Z_{\text{thresh.}}.
\end{align}
For a realistic fishery model, this threshold value could represent a population value beyond which drastic ecological consequences may ensue, e.g. a tipping point.
In our setting we have chosen $X_{\text{thresh.}}=Y_{\text{thresh.}}=Z_{\text{thresh.}}=0.05$ as a rule of thumb---given that under natural (unharvested) dynamics the populations range within values of 0.5 to 1, this would represent on the order of a 90-95% population decrease from its ‘‘natural’’ level.
This roughly represents a near-extinction event.

The reward function had two componenets:
First, at each time step the harvest $h_t$ was received as a reward.
Second, if the episode ended early (say, at time-step $t$) because of a near-extinction event, a negative reward of $-100/t$ was awarded.
This reward function balances the extractive motivation of the fishery with conservation goals that go beyond the scope of long-term sustainable harvests.

\textbf{Training a DRL agent.}
We trained a DRL agent parametrized by a neural network with two hidden, 64-neuron, layers, on a local server with 2 commercial GPUs.
We used the Ray framework\footnote{\url{https://docs.ray.io/}} for training, in particular we used a Ray PPOConfig object to build the algorithm using the default values for all hyper-parameters.
In particular, no hyperparameter tuning was performed.
The agent was trained for 300 iterations of the PPO optimization step.
The total training time was on the order of 30 minutes to 1 hour.

>>>>

\textbf{The Environment.}
As seen in Fig. **TBD CONCEPTUAL**, we consider four different models of the ecosystem with which the fishery interacts, each of this is called an *environment*, alluding to the environment in an RL setting.
To recap, these four environments are a single-species growth model \eqref{eq:may} for a harvested species; a three-species model \eqref{eq:3d model stoch} with a single harvested species; the same three-species model but with *two* harvested species; and, finally, a three-species model with a time-varying parameter and two harvested species.

The policies explored were functions from states $N_t$ to either a single number $M_X(N_t) = \pi_\theta(N_t)$, or a pair of numbers $(M_X(N_t), M_Y(N_t)) = \pi_\theta(N_t)$ depending on the environment considered.
\textbf{Probably a table showing the different combinations of species and fishery numbers would be good, possibly giving these ‘scenarios’ names.}

Our goal is to evaluate the performance of different policy strategies over a specified window of time.
We have chosen this window to be 200 timesteps, where each discrete time-step represents the dynamical evolution of the system over a year.
_Episodes_ begin at timestep zero and are controlled by the policy and the system's natural evolution until either the maximum timespan is reached, or until a ‘‘near-extinction event’’ occurs---that is, until any of the populations go below a given threshold,
\begin{align}
  \label{eq:thresh}
  X_t \leq X_{\text{thresh.}},\quad
  Y_t \leq Y_{\text{thresh.}},\quad
  \text{or,}\quad
  Z \leq Z_{\text{thresh.}}.
\end{align}
In our setting we have chosen $X_{\text{thresh.}}=Y_{\text{thresh.}}=Z_{\text{thresh.}}=0.05$ as a rule of thumb---given that under natural (unharvested) dynamics the populations range within values of 0.5 to 1, this would represent on the order of a 90-95% population decrease from its ‘‘natural’’ level.

The reward function defining our policy optimization problem has two componenets:
The first is economical: the total biomass harvested over an episode.
The second seeks to reflect conservation goals: if a near-extinction event occurs, the episode is ended early and a negative reward of $-100/t$ was awarded.
This reward function balances the extractive motivation of the fishery with conservation goals that go beyond the scope of long-term sustainable harvests.

\textbf{Training a DRL agent.}
We trained a DRL agent parametrized by a neural network with two hidden, 64-neuron, layers, on a local server with 2 commercial GPUs.
We used the Ray framework\footnote{\url{https://docs.ray.io/}} for training, in particular we used a Ray PPOConfig object to build the algorithm using the default values for all hyper-parameters.
In particular, no hyperparameter tuning was performed.
The agent was trained for 300 iterations of the PPO optimization step for the three-species cases.
The total training time was on the order of 30 minutes to 1 hour.
For the single-species model, the training iterations were scaled down to 100 and the training time was around 10 minutes.
